\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-10-05 demystify.tex
Updates from Overleaf

For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the U. VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf. Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters can be positive as they have helped many students getting into top PhD programs in the US.

@@

Note that funding is typically more available for PhD students than 
Masters. Undergraduate students typically have no funding and also have to pay international tuition, which is very expensive!.

@@

\subsection{How to rank or select a CS graduate program?}
   \begin{center}
    \includegraphics[scale=0.5]{c1.png}
   \end{center}

@@

@@

If you don't know the professor (e.g., first email contact), then use \textbf{Prof. Lastname} or \textbf{Dr. Lastname}. I've seen many international students write \textbf{Prof.} or \textbf{Dr.} \textbf{Firstname Lastname}.  Writing like that makes it like you copy and paste names, so no need to do so,  just Prof. or Dr. Lastname.
Moreover, if you don't know the prof., do not use Mr. or Mrs., or Firstname. To me it seems a bit disrespectful. As you know that prof. better and depends on their preference, you may call them by their firstname.  For example, my students and colleague call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.

@@

\paragraph{Acknowledgement} Many people have contributed to this document.
Profs. Craig Yu (GMU), Hakan Aydin (GMU), and Hung Le (UMass) provided valuable input in the early version. Other GMU faculty members also have provided feedback and contributions.  Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback. 
\textbf{Thank you!}

>>>>

For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the US. VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf. Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters can be positive as they have helped many students getting into top PhD programs in the US.

@@

Note that funding is typically more available for PhD students than 
Masters. Undergraduates typically have no funding and also have to pay international tuition, which is very expensive!.

@@

\section{Choosing Schools and Professors}
\subsection{Selecting and Ranking Schools for CS}
   \begin{center}
    \includegraphics[scale=0.5]{c1.png}
   \end{center}

@@

\begin{commentbox}[Xiaokuan]
Chinese students often only look at USNews rankings when selecting their Ph.D. universities (I did that, too, when I was applying for Ph.D. positions).
Now that I am a professor, I find it to be the least promising way.
%
The reason is that USNews does not provide a good metric for evaluating the quality of the Ph.D. program.
%
If you want to do great research, CSRankings is the best way to find good and active professors (which did not exist when I was applying),
since it solely focuses on publications at top-tier CS conferences.
%
Also, 
I think Ph.D. is not only about research; 
you need to also consider your daily life there, since you will (probably) stay for at least five years.
%
You might regret it if you did not consider this seriously before applying.
%
Whether your research interest matches that of your advisor is also important;
if there is a mismatch,
either you or your advisor has to make compromises,
which often leads to disagreements or conflicts.

\end{commentbox}

@@

If you don't know the professor (e.g., first email contact), then use \textbf{Prof. Lastname} or \textbf{Dr. Lastname}. I've seen many international students write \textbf{Prof.} or \textbf{Dr.} \textbf{Firstname Lastname}.  Writing like that makes it like you copy and paste names, so no need to do so,  just Prof. or Dr. Lastname.
Moreover, if you don't know the prof., do not use Mr. or Mrs., or Firstname---to me it seems a bit disrespectful. As you get to know that prof. better and depends on their preference, you may call them by their firstname.  For example, my students and colleague call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.

@@

\paragraph{Acknowledgement} Many people have contributed to this document.
Profs. Craig Yu (GMU), Hakan Aydin (GMU), 
Xiaokuan Zhang (GMU), and Hung Le (UMass) provided valuable input in the early version. Other GMU faculty members also have provided feedback and contributions.  Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback. 
\textbf{Thank you!}
